---
title: "Master File: How Proximity to National Parks Shapes Partisanship: Evidence from U.S. Counties, 2000-2024"
author: "Julianna Thomson and Stephen Gailliot"
date: "2026-02-16"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

**Notes:**

-   Chunk 1: Set up and import data.

-   Chunk 2: Clean and prepare data.

-   Chunk 3: Creates panel data.

-   Chunk 4: Adds Controls to panel data.

-   Chunk 5: Regression models for 50-mile radius around national parks.

-   Chunk 6: Line charts and stacked bar charts comparing counties and states.

-   Chunk 7: Regression based on proximity.

-   Chunk 8: State map for 2024 election only.

-   Chunk 9: State maps for all election years.

*Replication Rmd file contains the code to run abbreviated Chunks 5-9 using exported csv data.*

**Key Variables:**

*df_panel_sub_50* contains the panel data for the regressions.

-   DV is voteshare, which is the Democrat vote share by county.

-   IV1 is CnPdistance_log.
    This is the logged distance between the county and national park centroids.

-   IV2 is proximity.
    This is the logged sum of distances between the county and each national park within 50 miles by centroid.

-   Controls: WHITE for ethnicity, INCOME10K, POPULATION100K, BA for education (Bachelors or higher) and M1_distance_miles/M1_distance_log for distance to the nearest metropolitan area.

*model_50_summary* contains the results of the regression analysis based on df_panel_sub_50.
It accounts for counties that are up to 50 miles from a national park.

**Chunk 1:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#start new R session and clear output. 
rm(list = ls())
cat("\f")
set.seed(42)

#----SETTING UP----

Packages<-c("tidyr", "tidyverse", "dplyr","knitr", "readr", "sf", "units", "spdep", "ggplot2", "modelsummary", "stargazer", "gt", "magrittr")
lapply(Packages, library, character.only = TRUE)

#DATA IMPORT

setwd("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS")

#Read county-level voting data from MIT Election Data + Science Lab
#https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ
df_votes <- read_csv("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/countypres_2000-2024.csv")

#Read in geospatial data from the National Park Service (NPS).
#https://irma.nps.gov/DataStore/Reference/Profile/2224545?lnv=True
df_NPS <- st_read("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/nps_boundary.shp")

#Read in geospatial county data from the U.S. Census.
#https://www.census.gov/cgi-bin/geo/shapefiles/index.php
#Note: U.S. Census does not have a county file for 2004.
df_counties24 <- st_read("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/tl_2024_us_county.shp")

#Read in population estimates from U.S. Department of Agriculture Economic Research Service
#https://www.ers.usda.gov/data-products/county-level-data-sets/county-level-data-sets-download-data
df_pop <- read_csv("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/PopulationEstimates.csv")

#Read in median income from U.S. Department of Agriculture Economic Research Service
#https://www.ers.usda.gov/data-products/county-level-data-sets/county-level-data-sets-download-data
df_income <- read.csv("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/Unemployment2023.csv")

#Read in race/ethnicity data from U.S. Census Bureau
#https://data.census.gov/table/ACSDT5Y2023.B02001?t=Race+and+Ethnicity&g=010XX00US$0500000
df_race <- read.csv("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/ACSDT5Y2023.B02001-Data.csv")

#Read in education data from U.S. Census Bureau
#https://data.census.gov/table/ACSST5Y2023.S1501?t=Education&g=010XX00US$0500000
df_edu <- read.csv("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/ACSST5Y2023.S1501-Data.csv")

#Read in metropolitan/micropolitan data from U.S. Census
#https://catalog-beta.data.gov/dataset/tiger-line-shapefile-2018-nation-u-s-current-metropolitan-statistical-area-micropolitan-st
sf_CBSA <- st_read("C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/tl_2018_us_cbsa.shp")
```

**Chunk 2:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----CLEAN DATA----

#Calculate vote share for each candidate by county

df_votes$vote_share <- df_votes$candidatevotes / df_votes$totalvotes * 100

#Prepare voting data for merging with county data

#Duplicate the FIPS column and add a zero to each. Necessary to merge with the Census data below.
df_votes$county_fips2 <- as.character(df_votes$county_fips)
df_votes$county_fips2 <- sprintf("%05s", df_votes$county_fips2)
df_votes$GEOID <- df_votes$county_fips2

#Subset relevant NPS data

#Subset NPS data; National Park
df_parks <- subset(df_NPS, UNIT_TYPE=="National Park")

#5-digit GEOIDs for race/ethnicity data

df_race <- df_race[-1, ] #Remove header
df_race$GEOID <- stringr::str_sub(df_race$GEO_ID, -5)

#5-digit GEOIDs for education data

df_edu <- df_edu[-1, ] #Remove header
df_edu$GEO_ID <- stringr::str_sub(df_edu$GEO_ID, -5)

# ----CALCULATE COUNTY CENTROIDS----
#County Data

county_centroids <- st_centroid(df_counties24)
county_centroids <- cbind(
  county_centroids,
  st_coordinates(county_centroids))

#NPS Data

sf_use_s2(FALSE)  #Issue with overlap between two of the parks. This and the code that follows enabled centroid calculations.
parks_valid <- df_parks %>%
  st_make_valid()
parks_centroids <- parks_valid %>%
  st_centroid()

# ----MERGE CENTROID AND VOTING DATA----

df_votes_w_coords <- merge(df_votes, county_centroids, by.x="county_fips2", by.y="GEOID")

#----SET SPATIAL DATA TO EPSG 2163----
#EPSG 2163 is the US National Atlas Equal Area. 
parks_proj <- st_transform(parks_centroids, 2163) 
parks_proj$park_code <- make.unique(rep(LETTERS, length.out = nrow(parks_proj)))
parks_proj$park_code <- gsub("\\.", "", parks_proj$park_code)

counties_proj <- st_transform(county_centroids, 2163) 
counties_proj$county_code <- make.unique(rep(LETTERS, length.out = nrow(counties_proj)))
counties_proj$county_code <- gsub("\\.", "", counties_proj$county_code)

#----CALCULATE DISTANCE BETWEEN COUNTY AND PARK CENTROIDS----

#Create an index based on the parks and counties.
dist_mat <- st_distance(parks_proj, counties_proj)

#Conver the matrix to a dataframe.
df_dist_CnP <- as.data.frame(as.table(dist_mat))
colnames(df_dist_CnP) <- c("park_code", "county_code", "distance_m")
df_dist_CnP <- as.data.frame(as.table(dist_mat))

rownames(dist_mat) <- parks_proj$park_code
colnames(dist_mat) <- counties_proj$county_code 

df_dist_CnP <- as.data.frame(as.table(dist_mat)) %>%
  rename(park_code = Var1, county_code = Var2, distance_m = Freq)

#Join the dataframe with distances between parks and counties by centroid with the full park and county dataframes. This brings back in the spatial data and identifiers (i.e., GEOID)
df_dist_CnP <- df_dist_CnP %>%
  left_join(parks_proj %>% st_set_geometry(NULL), by = "park_code") %>%
  left_join(counties_proj %>% st_set_geometry(NULL), by = "county_code")

#Add a column for miles, as the default is meters.
df_dist_CnP <- df_dist_CnP %>%
  mutate(distance_miles = set_units(distance_m, "miles") %>% drop_units())

#----CALCULATE DISTANCES, COUNTIES AND METROPOLITAN AREAS------

#df_dist_CnP_50 <- subset(df_dist_CnP, distance_miles <=50)

cbsa_m1_proj <- sf_CBSA %>%
  filter(LSAD == "M1")
cbsa_m1_proj <- st_centroid(cbsa_m1_proj)
cbsa_m1_proj <- st_transform(cbsa_m1_proj, 2163)
cbsa_m1_proj$M1_code <- make.unique(rep(LETTERS, length.out = nrow(cbsa_m1_proj)))
cbsa_m1_proj$M1_code <- gsub("\\.", "", cbsa_m1_proj$M1_code)

#Create an index based on the M1s and counties.
dist_mat2 <- st_distance(cbsa_m1_proj, counties_proj)

#Conver the matrix to a dataframe.
df_dist_CnM1 <- as.data.frame(as.table(dist_mat2))
colnames(df_dist_CnM1) <- c("M1_code", "county_code", "distance_m")
df_dist_CnM1 <- as.data.frame(as.table(dist_mat2))

rownames(dist_mat2) <- cbsa_m1_proj$M1_code
colnames(dist_mat2) <- counties_proj$county_code 

df_dist_CnM1 <- as.data.frame(as.table(dist_mat2)) %>%
  rename(M1_code = Var1, county_code = Var2, distance_m = Freq)

#Join the dataframe with distances between M1s and counties by centroid. This brings back in the spatial data and identifiers (i.e., GEOID)
df_dist_CnM1 <- df_dist_CnM1 %>%
  left_join(cbsa_m1_proj %>% st_set_geometry(NULL), by = "M1_code") %>%
  left_join(counties_proj %>% st_set_geometry(NULL), by = "county_code")

#Add a column for miles, as the default is meters.
df_dist_CnM1 <- df_dist_CnM1 %>%
  mutate(distance_miles = set_units(distance_m, "miles") %>% drop_units())
```

**Chunk 3:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----CREATE PANEL DATA----

df_dem <- subset(df_votes, party=="DEMOCRAT")

df_dem_votes <- df_dem %>%
  group_by(GEOID, year) %>%
  summarise(
    candidatevotes = sum(candidatevotes, na.rm = TRUE),
    .groups = "drop"
  )

df_dem_total <- df_dem %>%
  group_by(GEOID, year) %>%
  summarise(
    totalvotes = sum(totalvotes, na.rm = TRUE),
    .groups = "drop"
  )

df_dem_voteshare <- df_dem_votes
df_dem_voteshare$totalvotes <- df_dem_total$totalvotes
df_dem_voteshare$voteshare <- df_dem_voteshare$candidatevotes / df_dem_voteshare$totalvotes * 100

df_dem_voteshare_dist <- df_dem_voteshare %>%
  left_join(df_dist_CnP, by = "GEOID")

df_dem_voteshare_dist$distanceCnP_log <- log(df_dem_voteshare_dist$distance_miles)

#Save panel data as csv for replication file (need it to create state maps for multiple election years.)
write_csv(df_dem_voteshare, "C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/df_dem_voteshare.csv")
```

**Chunk 4:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----CONTROLS----
#------Population------
df_pop_23 <- subset(df_pop, Attribute=="POP_ESTIMATE_2023")
us_states_territories <- c(
  "United States", "Alabama","Alaska","Arizona","Arkansas","California","Colorado",
  "Connecticut","Delaware","Florida","Georgia","Hawaii","Idaho",
  "Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana",
  "Maine","Maryland","Massachusetts","Michigan","Minnesota",
  "Mississippi","Missouri","Montana","Nebraska","Nevada",
  "New Hampshire","New Jersey","New Mexico","New York",
  "North Carolina","North Dakota","Ohio","Oklahoma","Oregon",
  "Pennsylvania","Rhode Island","South Carolina","South Dakota",
  "Tennessee","Texas","Utah","Vermont","Virginia","Washington",
  "West Virginia","Wisconsin","Wyoming",
  "District of Columbia",
  "Puerto Rico","Guam","American Samoa",
  "Northern Mariana Islands","U.S. Virgin Islands"
)

df_pop_23 <- df_pop_23[!df_pop_23$Area_Name %in% us_states_territories, ]
df_pop_23 <- df_pop_23[, c(1, ncol(df_pop_23))]
colnames(df_pop_23) <- c("GEOID", "POPULATION")
df_pop_23$GEOID <- as.numeric(df_pop_23$GEOID)
df_pop_23$GEOID <- stringr::str_pad(df_pop_23$GEOID, width = 5, side = "left", pad = "0")
df_pop_23$POPULATION100K <- df_pop_23$POPULATION / 100000

#------Median Income------
df_income_23 <- subset(df_income, Attribute=="Median_Household_Income_2022")
df_income_23 <- df_income_23[!df_income_23$Area_Name %in% us_states_territories, ]
df_income_23 <- df_income_23[, c(1, ncol(df_income_23))]
colnames(df_income_23) <- c("GEOID", "INCOME")
df_income_23$GEOID <- as.numeric(df_income_23$GEOID)
df_income_23$GEOID <- stringr::str_pad(df_income_23$GEOID, width = 5, side = "left", pad = "0")
df_income_23$INCOME <- as.numeric(df_income_23$INCOME)
df_income_23$INCOME10K <- df_income_23$INCOME / 10000

#------Race/Ethnicity------
#Calculate the percent of the white population for each county. Total population is in B02001_001E. Total white population is in B02001_002E
df_race$B02001_001E <- as.numeric(df_race$B02001_001E)
df_race$B02001_002E <- as.numeric(df_race$B02001_002E)
df_race$WHITE <- df_race$B02001_002E / df_race$B02001_001E * 100
df_race_W <- df_race[, (ncol(df_race) - 1):ncol(df_race)]

#------Education-----
df_race$B02001_001E <- as.numeric(df_race$B02001_001E)
df_race$B02001_002E <- as.numeric(df_race$B02001_002E)

#Sum all populations
#18-24: S1501_C01_001E
#25+: S1501_C01_006E
#18-24 BA or higher: S1501_C01_005E
#25+ BA or higher: S1501_C01_012E

df_edu$S1501_C01_001E <- as.numeric(df_edu$S1501_C01_001E)
df_edu$S1501_C01_006E <- as.numeric(df_edu$S1501_C01_006E)
df_edu$S1501_C01_005E <- as.numeric(df_edu$S1501_C01_005E)
df_edu$S1501_C01_012E <- as.numeric(df_edu$S1501_C01_012E)

df_edu$allpop <- df_edu$S1501_C01_001E + df_edu$S1501_C01_006E
df_edu$BA <- df_edu$S1501_C01_005E + df_edu$S1501_C01_012E
df_edu$deg <- df_edu$BA / df_edu$allpop * 100

df_BA <- df_edu[, c(1, ncol(df_edu))]
colnames(df_BA) <- c("GEOID", "BA")

#------Nearest Metropolitan Area (M1)------

df_dist_CnM1_min <- df_dist_CnM1 %>%
  group_by(GEOID.y) %>%
  summarise(M1_distance_miles = min(distance_miles, na.rm = TRUE), .groups = "drop") 

#----ADD CONTROLS TO PANEL DATA----

df_panel <- df_dem_voteshare_dist %>%
  left_join(df_race_W, by = "GEOID")

df_panel <- df_panel %>%
  left_join(df_income_23, by = "GEOID")

df_panel <- df_panel %>%
  left_join(df_pop_23, by = "GEOID")

df_panel <- df_panel %>%
  left_join(df_BA, by = "GEOID")

df_panel <- df_panel %>%
  left_join(df_dist_CnM1_min, by = c("GEOID" = "GEOID.y"))

df_panel_sub <- data.frame(
  GEOID = df_panel$GEOID,
  year = df_panel$year,
  voteshare = df_panel$voteshare,
  CnP_distance_miles = df_panel$distance_miles,
  CnP_distance_log = df_panel$distanceCnP_log,
  M1_distance_miles = df_panel$M1_distance_miles,
  WHITE = df_panel$WHITE,
  INCOME10K = df_panel$INCOME10K,
  POPULATION100K = df_panel$POPULATION100K,
  BA = df_panel$BA
)

df_panel_sub <- na.omit(df_panel_sub)

df_panel_sub$M1_distance_log <- log(df_panel_sub$M1_distance_miles)
```

**Chunk 5:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----REGRESSION MODELS----

#Regression: Logged Distance (Miles)

df_panel_sub_50 <- subset(df_panel_sub, CnP_distance_miles <=50)

#Export as csv for replication file to run these first regressions. 
write_csv(df_panel_sub_50, "C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/df_panel_sub_50.csv")

model_50_log <- lm(
  voteshare ~ CnP_distance_log
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50
)

#------Print Model Summary------
model_50_log_summary <- modelsummary(
  list("Model" = model_50_log),

  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_50_log_summary

#------Regression: Logged Distance (Miles) with Factor(year)

model_50_fy_log <- lm(
  voteshare ~ CnP_distance_log + factor(year)
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50
)

#------Print Model Summary------
model_50_fy_log_summary <- modelsummary(
  list("Model" = model_50_fy_log),

  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_50_fy_log_summary
```

**Chunk 6:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----COMPARISON: DEMOCRAT VOTE SHARE, PARK PROXIMITY TO STATE----

#Calculate total Democrat voteshare by state and year

df_state_votes <- df_dem %>%
  group_by(state, year) %>%
  summarise(state_total = sum(totalvotes, na.rm=TRUE),
            .groups = "drop")

df_state_votes_dem <- df_dem %>%
  group_by(state, year) %>%
  summarise(dem_total = sum(candidatevotes, na.rm=TRUE),
            .groups = "drop")

df_state_votes_dem$votes <- df_state_votes$state_total

df_state_votes_dem$dem_vote_share <- (df_state_votes_dem$dem_total / df_state_votes_dem$votes) * 100

df_compare_vote_share <- df_state_votes_dem %>%
  left_join(df_dem, by = c("state","year"))

df_compare_vote_share <- df_compare_vote_share %>%
   semi_join(df_panel_sub_50, by = c("GEOID", "year"))

df_compare_vote_share$vs_compare <- df_compare_vote_share$dem_vote_share - df_compare_vote_share$vote_share

df_compare_vote_share_clean <- df_compare_vote_share %>%
  group_by(state, year) %>%
  summarise(
    vs_compare = mean(vs_compare, na.rm = TRUE),
    .groups = "drop"
  )

#Again, exporting data needed to create plots as CSV to run replication file.
write_csv(df_compare_vote_share_clean, "C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/df_compare_vote_share_clean.csv")

plot_compare <- ggplot(df_compare_vote_share_clean,
                       aes(x = year, y = vs_compare, group = state)) +
  geom_line(color = "gray70", alpha = 0.3, linewidth = 0.4) +
  
  # Add overall mean trend
  stat_summary(aes(group = 1),
               fun = mean,
               geom = "line",
               color = "blue",
               linewidth = 1.2) +
  
  scale_x_continuous(
    breaks = seq(min(df_compare_vote_share_clean$year),
                 max(df_compare_vote_share_clean$year),
                 by = 4)
  ) +
  
  coord_cartesian(ylim = c(-50, 50)) +
  theme_minimal(base_size = 12) +
  labs(
    title = "State–County Vote Share Difference Over Time",
    subtitle = "Gray lines = states | Blue line = average",
    x = "Year",
    y = "Vote Share Difference (%)"
  )

plot_compare2<-ggplot(df_compare_vote_share_clean,
       aes(x = year, y = vs_compare)) +
  geom_line(linewidth = 0.6, color = "steelblue") +
  facet_wrap(~ state) +
  theme_minimal() +
  labs(
    title = "Vote Share Difference by State Over Time",
    x = "Year",
    y = "Vote Share Difference (%)"
  )

plot_compare3<-ggplot(df_compare_vote_share_clean,
       aes(x = year, y = reorder(state, vs_compare), fill = vs_compare)) +
  geom_tile() +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "blue",
                       midpoint = 0) +
  theme_minimal() +
  labs(
    title = "State–County Vote Share Difference Over Time",
    x = "Year",
    y = "State",
    fill = "Vote Share\nDifference"
  )
```

**Chunk 7:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----PROXIMITY ANALYSIS----

#Calculate total proximity for bases within 50 miles.

df_panel_sub_50 <- df_panel_sub %>%
  filter(CnP_distance_miles <= 50, year == 2024)

df_panel_sub_50 <- df_panel_sub_50 %>%
  mutate(proximity = 50 - CnP_distance_miles)

df_panel_sub_50_prox <- df_panel_sub_50 %>%
  group_by(GEOID) %>%
  summarise(total_proximity = sum(proximity, na.rm=TRUE),
            .groups = "drop")

df_panel_sub_50$proximity <- df_panel_sub_50$proximity[match(df_panel_sub_50$GEOID, df_panel_sub_50$GEOID)]

#Once df_panel_sub_50 is overwritten, export under a new name to run replication files: 
write_csv(df_panel_sub_50, "C:/Users/julia/OneDrive - George Mason University - O365 Production/Stephen G Gailliot's files - NPS/df_panel_sub_50_prox.csv")

#Proximity regression

model_prox <- lm(
  voteshare ~ log(proximity)
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50
)

#------Print Model Summary------
model_prox_summary <- modelsummary(
  list("Model" = model_prox),
  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_prox_summary
```

**Chunk 8:**

```{r, message=FALSE, warning=FALSE, results='hide'}
#----STATE MAP for 2024 Election only
# Filter counties to Wyoming 
wy_counties <- df_counties24 %>%
  filter(STATEFP == "56")

# Filter NPS data to National Parks only
#unique(df_NPS$UNIT_NAME) #to get park names
yellowstone <- df_NPS %>%
  filter(UNIT_NAME == "Yellowstone National Park")
grandteton <- df_NPS %>%
  filter(UNIT_NAME == "Grand Teton National Park")

#You could also do it this way, however, Yellowstone is classified as an Idaho park so this won't pull it into Wyoming.
#seems like we'll have a similar problem with Gateway Arch NP. 
#wy_parks <- df_NPS %>%
  #filter(STATE == "WY",
         #UNIT_TYPE == "National Park")

# Make sure both layers use the same coordinate reference system
yellowstone <- st_transform(yellowstone, st_crs(wy_counties))
grandteton <- st_transform(grandteton, st_crs(wy_counties))

#Filter voting data to Wyoming, 2024, Democrat
wy_dem_2024 <- df_votes %>%
  filter(state_po == "WY",
         year == 2024,
         party == "DEMOCRAT")

# Merge with county shapefile
wy_counties <- wy_counties %>%
  left_join(wy_dem_2024, by = "GEOID")

# Make Map
ggplot() +
  geom_sf(data = wy_counties, aes(fill = vote_share), color = "black", linewidth = 0.5) +
  geom_sf(data = yellowstone, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.8) +
  geom_sf(data = grandteton, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.8) +
  scale_fill_gradient(
    low = "white", high = "royalblue4",
    name = "Dem. Vote Share (%)",
    limits = c(0, 100)
  ) +
  labs(
    title = "2024 Democrat Vote Share by County - Wyoming",
    subtitle = "With Yellowstone National Park and Grand Teton National Park",
    caption = "Sources: MIT Election Lab, U.S. Census Bureau, National Park Service"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    plot.caption = element_text(hjust = 0.5, size = 8),
    legend.position = "bottom"
  )

```

**Chunk 9:**

```{r}
#----STATE MAP for all Election Years

# Filter counties to Wyoming 
wy_counties <- df_counties24 %>%
  filter(STATEFP == "56")

# Filter NPS data to National Parks only
yellowstone <- df_NPS %>%
  filter(UNIT_NAME == "Yellowstone National Park")
grandteton <- df_NPS %>%
  filter(UNIT_NAME == "Grand Teton National Park")

# Match CRS
yellowstone <- st_transform(yellowstone, st_crs(wy_counties))
grandteton  <- st_transform(grandteton,  st_crs(wy_counties))

# Filter vote share data to Wyoming across all election years
wy_voteshare_allyears <- df_dem_voteshare %>%
  filter(GEOID %in% wy_counties$GEOID,
         year %in% c(2000, 2004, 2008, 2012, 2016, 2020, 2024))

# Join with county shapefile (one row per county per year)
wy_counties_allyears <- wy_counties %>%
  left_join(wy_voteshare_allyears, by = "GEOID")

# Make faceted map

ggplot() +
  geom_sf(data = wy_counties_allyears, aes(fill = voteshare), color = "black", linewidth = 0.3) +
  geom_sf(data = yellowstone, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.6) +
  geom_sf(data = grandteton, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.6) +
  scale_fill_gradient(
    low = "white", high = "royalblue4",
    name = "Dem. Vote Share (%)",
    limits = c(0, 100)
  ) +
  facet_wrap(~year, ncol = 4) +
  labs(
    title = "Democrat Vote Share by County - Wyoming (2000–2024)",
    subtitle = "Green shading indicates Yellowstone and Grand Teton National Parks",
    caption = "Sources: MIT Election Lab, U.S. Census Bureau, National Park Service"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    plot.caption  = element_text(hjust = 0.5, size = 8),
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 10)
  )

```
