---
title: "Replication File for: How Proximity to National Parks Shapes Partisanship: Evidence from U.S. Counties, 2000-2024"
author: "Julianna Thomson and Stephen Gailliot"
date: "2026-02-26"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

First, load packages, and adjust path files to your machine.
You'll need to read in these files to run the analyses:

-   df_panel_sub_50.csv

-   df_compare_vote_share_clean.csv

-   df_panel_sub_50_prox.csv

-   df_dem_voteshare.csv

-   nps_boundary.shp

-   tl_2024_us_county.shp

```{r, message=FALSE, warning=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
cat("\f")
set.seed(42)

# Load Packages and set your working directory

Packages<-c("tidyr", "tidyverse", "dplyr","knitr", "readr", "sf", "units", "spdep", "ggplot2", "modelsummary", "stargazer", "gt", "magrittr")
lapply(Packages, library, character.only = TRUE)

df_panel_sub_50 <- read_csv("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/df_panel_sub_50.csv")

df_compare_vote_share_clean <- read_csv("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/df_compare_vote_share_clean.csv")

df_panel_sub_50_prox <- read_csv("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/df_panel_sub_50_prox.csv")

df_dem_voteshare <- read_csv("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/df_dem_voteshare.csv")

#Read in geospatial data from the National Park Service (NPS).
#https://irma.nps.gov/DataStore/Reference/Profile/2224545?lnv=True
df_NPS <- st_read("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/nps_boundary.shp")

#Read in geospatial county data from the U.S. Census.
#https://www.census.gov/cgi-bin/geo/shapefiles/index.php
#Note: U.S. Census does not have a county file for 2004.
df_counties24 <- st_read("~/Library/CloudStorage/OneDrive-GeorgeMasonUniversity-O365Production/Stephen G Gailliot's files - NPS/tl_2024_us_county.shp")
```

This code chunk runs the first set of regression analyses.

-   Regression 1: Logged Distance (Miles)

-   Regression 2: Logged Distance (Miles) with Factor(year)

*df_panel_sub_50* contains the panel data for the regressions:

-   DV is voteshare, which is the Democrat vote share by county.

-   IV1 is CnPdistance_log.
    This is the logged distance between the county and national park centroids.

-   IV2 is proximity.
    This is the logged sum of distances between the county and each national park within 50 miles by centroid.

-   Controls: WHITE for ethnicity, INCOME10K, POPULATION100K, BA for education (Bachelors or higher) and M1_distance_miles/M1_distance_log for distance to the nearest metropolitan area.

*model_50_summary* contains the results of the regression analysis based on df_panel_sub_50.
It accounts for counties that are up to 50 miles from a national park.

```{r}
#----Regression 1: Logged Distance (Miles)----

model_50_log <- lm(
  voteshare ~ CnP_distance_log
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50
)

#------Print Model Summary------
model_50_log_summary <- modelsummary(
  list("Model" = model_50_log),

  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_50_log_summary

#----Regression 2: Logged Distance (Miles) with Factor(year)----

model_50_fy_log <- lm(
  voteshare ~ CnP_distance_log + factor(year)
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50
)

#------Print Model Summary------
model_50_fy_log_summary <- modelsummary(
  list("Model" = model_50_fy_log),

  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_50_fy_log_summary
```

This code chunk runs the second analysis.
This is a comparison of Democrat vote share and park proximity to state vote share.
This code chunk generates three plots.

```{r}
#----Plot 1: State-County Vote Share Difference Over Time----
plot_compare <- ggplot(df_compare_vote_share_clean,
                       aes(x = year, y = vs_compare, group = state)) +
  geom_line(color = "gray70", alpha = 0.3, linewidth = 0.4) +
  
  # Add overall mean trend
  stat_summary(aes(group = 1),
               fun = mean,
               geom = "line",
               color = "blue",
               linewidth = 1.2) +
  
  scale_x_continuous(
    breaks = seq(min(df_compare_vote_share_clean$year),
                 max(df_compare_vote_share_clean$year),
                 by = 4)
  ) +
  
  coord_cartesian(ylim = c(-50, 50)) +
  theme_minimal(base_size = 12) +
  labs(
    title = "State–County Vote Share Difference Over Time",
    subtitle = "Gray lines = states | Blue line = average",
    x = "Year",
    y = "Vote Share Difference (%)"
  )

#----Plot 2: Vote Share Difference by State Over Time----
plot_compare2<-ggplot(df_compare_vote_share_clean,
       aes(x = year, y = vs_compare)) +
  geom_line(linewidth = 0.6, color = "steelblue") +
  facet_wrap(~ state) +
  theme_minimal() +
  labs(
    title = "Vote Share Difference by State Over Time",
    x = "Year",
    y = "Vote Share Difference (%)"
  )

#----Plot 3: State-County Vote Share Difference Over Time----
plot_compare3<-ggplot(df_compare_vote_share_clean,
       aes(x = year, y = reorder(state, vs_compare), fill = vs_compare)) +
  geom_tile() +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "blue",
                       midpoint = 0) +
  theme_minimal() +
  labs(
    title = "State–County Vote Share Difference Over Time",
    x = "Year",
    y = "State",
    fill = "Vote Share\nDifference"
  )
```

This last chunk of code runs the final proximity regression.

```{r}
model_prox <- lm(
  voteshare ~ log(proximity)
  + WHITE + INCOME10K + POPULATION100K + BA + M1_distance_log,
  data = df_panel_sub_50_prox
)


#------Print Model Summary------
model_prox_summary <- modelsummary(
  list("Model" = model_prox),
  stars = TRUE,
  statistic = "({std.error})",
  gof_omit = "IC|Log|Adj",
  title = "Regression Results",
  output = "gt"
  ) %>%
  tab_options(
    data_row.padding = px(1),
    heading.padding = px(2),
    table.font.size = px(11)
)

model_prox_summary 

```

Lastly, to generate the maps for each election year, use the code below.
Once you read in the necessary data, you can adjust this code to make a map of any state/county/park you're interested in.

Use this link to look up the state FIPS code: <https://www.bls.gov/respondents/mwr/electronic-data-interchange/appendix-d-usps-state-abbreviations-and-fips-codes.htm>

```{r}
# Next, filter counties to the state you are interested in by adjusting the FIPS code.   
wy_counties <- df_counties24 %>%
  filter(STATEFP == "56")

# Filter NPS data to National Parks in Wyoming only
yellowstone <- df_NPS %>%
  filter(UNIT_NAME == "Yellowstone National Park")
grandteton <- df_NPS %>%
  filter(UNIT_NAME == "Grand Teton National Park")

#Match CRS
yellowstone <- st_transform(yellowstone, st_crs(wy_counties))
grandteton  <- st_transform(grandteton,  st_crs(wy_counties))

#Filter vote share data to Wyoming across all presidential election years
wy_voteshare_allyears <- df_dem_voteshare %>%
  filter(GEOID %in% wy_counties$GEOID,
         year %in% c(2000, 2004, 2008, 2012, 2016, 2020, 2024))

# Join with county shapefile (one row per county per year)
wy_voteshare_allyears$GEOID <- as.character(wy_voteshare_allyears$GEOID)
wy_counties$GEOID <- as.character(wy_counties$GEOID)
wy_counties_allyears <- wy_counties %>%
  left_join(wy_voteshare_allyears, by = "GEOID")

#Plot
ggplot() +
  geom_sf(data = wy_counties_allyears, aes(fill = voteshare), color = "black", linewidth = 0.3) +
  geom_sf(data = yellowstone, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.6) +
  geom_sf(data = grandteton, fill = "forestgreen", color = "darkgreen",
          alpha = 0.5, linewidth = 0.6) +
  scale_fill_gradient(
    low = "white", high = "royalblue4",
    name = "Dem. Vote Share (%)",
    limits = c(0, 100)
  ) +
  facet_wrap(~year, ncol = 4) +
  labs(
    title = "Democrat Vote Share by County - Wyoming (2000–2024)",
    subtitle = "Green shading indicates Yellowstone and Grand Teton National Parks",
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    plot.caption  = element_text(hjust = 0.5, size = 8),
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 10)
  )

```
